Laplacian envelope

### 2015-05-27 Notes/laplace.tex
laplace envelope notes

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{?.tex} with its contents
%       or include ?.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%equilibrium distribution
\newcommand{\eq}{\mathbf{p}^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\uv}{\mathbf{u}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\enc}{\mathbf{q}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\F}{\boldsymbol{\Phi}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\env}{Env}
\newcommand{\rh}{\hat{r}}
\newcommand{\CI}{\mathcal{I}}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
% ----------------------------------------------------------------
\input{sl_theorems_preamble.tex}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Area maximisation in continuous time}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

@@

\begin{abstract}
  We try to find the continuous time Markov process that has the maximal area under the signal-to-noise curve.
\end{abstract}

@@

\section{Continuous time Markov processes}\label{sec:ContMarkov}

In this section we'll provide a summary of all the relevant properties of ergodic Markov chains in continuous time.
It is a straightforward generalisation of material that can be found in \cite{kemeny1960finite} with some ideas from \cite{hunter2000survey}.

\subsection{Notation}\label{sec:not}

For any matrix $\mathbf{A}$, we define matrices $\mathbf{A}\dg$ and $\overline{\mathbf{A}}$ as
%
\begin{equation}\label{eq:dgdef}
  \mathbf{A}\dg_{ij} \equiv \delta_{ij}\mathbf{A}_{ij},
  \qquad
  \overline{\mathbf{A}} \equiv \mathbf{A}-\mathbf{A}\dg.
\end{equation}
%
We let $\onev$ denote a column-vector of ones and $\onem=\onev\onev\trans$ denote a matrix of ones.

A Markov process is described by a matrix of transitions rates, $\MM_{ij}$, from state $i$ to $j$.
The probabilities of being in each state at time $t$, the row-vector $\mathbf{p}(t)$, evolve according to
%
\begin{equation}\label{eq:rowsum}
  \diff{\mathbf{p}(t)}{t} = \mathbf{p}(t) \MM,
  \qquad
  \mathbf{p}(t)\onev=1,
  \qquad
  \MM\onev=0.
\end{equation}
%
In Surya's notes, $\MM=r\frg$.

The equilibrium probabilities, $\eq$, satisfy
%
\begin{equation}\label{eq:equilibrium}
  \eq\MM=0,
  \qquad
  \eq\onev=1.
\end{equation}
%
As we assume an ergodic process, this eigenvalue is non-degenerate.
If all other eigenvalues have strictly negative real parts, the process is regular (aperiodic).

We define additional matrices
%
\begin{equation}\label{eq:defDLP}
  \Lb \equiv (-\MM\dg)\inv,
  \qquad
  \Pb \equiv \I + \Lb\MM.
\end{equation}
%
It can be shown that $\Lb_{ii}$ is the mean time it takes to leave state $i$ and $\Pb_{ij}$ is the probability the the next transition from state $i$ goes to state $j$:
%
\begin{equation}\label{eq:LamdaPcmpt}
  \Lb_{ii} = \frac{1}{\sum_{j \neq j} \MM_{ij}},
  \qquad
  \Pb_{ij} = %\left\{
%  \begin{aligned}
%     &0                                         && \text{if }i=j, \\
%     &\frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &&\text{otherwise}.
%  \end{aligned}
%  \right.
  \begin{cases}
     0                                         &\text{if }i=j, \\
     \frac{\MM_{ij}}{\sum_{k \neq j} \MM_{ik}} &\text{otherwise}.
  \end{cases}
\end{equation}
%
Furthermore, we also define
%
\begin{equation}\label{eq:pdotD}
  \D \equiv \operatorname{diag}(\eq)\inv,
  \qquad\implies\qquad
  \eq\D=\onev\trans.
\end{equation}
%

\subsection{Fundamental matrix}\label{sec:fund}

\begin{defn}[Fundamental matrix]
  %
  \begin{equation}\label{eq:funddef}
    \fund \equiv (-\MM + \onev\pib)\inv,
  \end{equation}
  %
  where $\pib$ is any row-vector with $\pib\onev=1/\tau\neq0$.
\end{defn}
Note that the canonical choice for the discrete time version, $\pib=\eq$, is not available here due to problems with units.
It will be helpful to choose $\pib$ to be independent of $\MM$, \eg $\pib=\onev\trans/(n\tau)$.
All quantities that we calculate using $\fund$ below will be independent of this choice.

\begin{thm}
  The definition of $\fund$ is valid, \ie $(-\MM + \onev\pib)$ is invertible.
\end{thm}
\begin{proof}
  Assume there exists an $\mathbf{x}$ such that
  %
  \begin{equation}\label{eq:fundinvkern}
    (-\MM + \onev\pib)\mathbf{x}=0.
  \end{equation}
  %
  Multiplying from the left with $\eq$ gives
  %
  \begin{equation}\label{eq:fundinvpix}
    \pib \mathbf{x} = 0.
  \end{equation}
  %
  Substituting back into \eqref{eq:fundinvkern} gives
  %
  \begin{equation*}
    \MM\mathbf{x}=0.
  \end{equation*}
  %
  As we assume an ergodic process, the zero eigenvalue is non-degenerate.
  Therefore, $\mathbf{x}=\lambda\onev$.
  Substituting this into \eqref{eq:fundinvpix} gives
  %
  \begin{equation*}
    \lambda\pib\onev = \frac{\lambda}{\tau} = 0.
  \end{equation*}
  %
  As we defined $\pib$ such that $1/\tau\neq0$, this means $\lambda=0 \implies \mathbf{x}=0$.
\end{proof}

\begin{cor}
  %
  \begin{align}
    \pib\fund &= \eq, \label{eq:fundprob}\\
    \fund\onev &= \tau\onev,\label{eq:fundrowsum}\\
    \I+\MM\fund &= \onev\eq, \label{eq:fundQZ}\\
    \I+\fund\MM &= \tau\onev\pib. \label{eq:fundZQ}
  \end{align}
  %
\end{cor}
\begin{proof}
  We can deduce \eqref{eq:fundprob} and \eqref{eq:fundrowsum} be pre/post-multiplying the following equations by $\fund$:
  %
  \begin{equation*}
    \begin{aligned}
      \eq(-\MM + \onev\pib) &= \pib, \\
      (-\MM + \onev\pib)\onev &= \frac{\onev}{\tau}.
    \end{aligned}
  \end{equation*}
  %
  We can then deduce \eqref{eq:fundQZ} and \eqref{eq:fundZQ} by substituting these into
  %
  \begin{equation*}
    (-\MM + \onev\pib)\fund = \fund(-\MM + \onev\pib) = \I.
  \end{equation*}
  %
\end{proof}

%We will not need to use the concept of a g-inverse anywhere, so you can skip ahead to \sref{sec:fpt} now. For the sake of completeness, we can define $\MM^\#$ as
%%
%\begin{equation}\label{eq:ginvdef}
%  \MM^\# \equiv \tau\onev\eq - \fund
%\end{equation}
%%
%\begin{thm}
%  The matrix $\MM^\#$ satisfies the following defining properties of a g-inverse of $\MM$:
%  %
%  \begin{align}
%    \MM^\# \MM \MM^\# &= \MM^\#,     \label{eq:ginvpropo} \\
%    \MM \MM^\# \MM    &= \MM,        \label{eq:ginvpropi} \\
%    \MM \MM^\#        &= \MM^\# \MM, \label{eq:ginvcomm}
%  \end{align}
%  %
%  with \eqref{eq:ginvcomm} holding iff $\eq$ and $\pib$ are proportional to each other.
%\end{thm}
%\begin{proof}
%  Using \eqref{eq:rowsum}, \eqref{eq:equilibrium}, \eqref{eq:fundprob} and \eqref{eq:fundrowsum}, one can verify
%  %
%  \begin{alignat*}{2}
%    \MM \MM^\# &= -\MM \fund,
%    &\qquad
%    \pib \MM^\# &= 0,
%    \\
%    \MM^\# \MM &= -\fund \MM,
%    &
%    \MM^\# \onev &= 0.
%  \end{alignat*}
%  %
%  Substituting into \eqref{eq:fundQZ} and \eqref{eq:fundZQ} leads to
%  %
%  \begin{equation*}
%    \I - \MM \MM^\# = \onev \eq,
%    \qquad
%    \I - \MM^\# \MM = \tau \onev \pib.
%  \end{equation*}
%  %
%  Pre/post-multiplying these by $\MM$ or $\MM^\#$ produces \eqref{eq:ginvpropo} and \eqref{eq:ginvpropi}. Subtracting produces
%  %
%  \begin{equation*}
%    [\MM,\MM^\#] = \onev(\tau\pib-\eq),
%  \end{equation*}
%  %
%  which vanishes iff $\eq$ and $\pib$ are proportional to each other (due to $\pib\onev=1/\tau$ and $\eq\onev=1$, the constant of proportionality could only be $\tau$).
%\end{proof}

\subsection{First passage times}\label{sec:fpt}

\begin{defn}[First passage time matrix]
 We define $\fptb_{ij}$ as the mean time it takes the process to reach state $j$ for the first time, starting from state $i$.
 We also define $\fpt\dg_{ii}$ as the mean time it takes the process to return to state $i$.
 As usual, $\fpt=\fptb+\fpt\dg$.
\end{defn}

Consider the first time the process leaves state $i$.
On average, this will take time $\Lb_{ii}$.
With probability $\Pb_{ij}$, it will go directly to $j$, so the conditional mean time would be $\Lb_{ii}$.
On the other hand, if it goes to some other state, $k$, with probability $\Pb_{ik}$, the conditional mean time would be $\fptb_{kj}+\Lb_{ii}$.
Combining these, we get the recursion relation
%
\begin{equation}\label{eq:fptrecursion}
  \begin{aligned}
  \fpt_{ij} &= \sum_{k\neq j} \Pb_{ik}(\fptb_{kj}+\Lb_{ii}) + \Pb_{ij}\Lb_{ii} \\
    &= \sum_{k\neq j} \Pb_{ik}\fptb_{kj} + \sum_{k} \Pb_{ik}\Lb_{ii}\\
    &= \sum_{k\neq j} \Pb_{ik}\fptb_{kj} + \Lb_{ii},\\
   \fpt &= \Pb \fptb + \Lb\onem,
  \end{aligned}
\end{equation}
%
or, using \eqref{eq:defDLP},
%
\begin{equation}\label{eq:fptrecdg}
  \fpt\dg = \Lb\MM\fptb+\Lb\onem.
\end{equation}
%

\begin{thm}
  The recurrence times are given by
  %
  \begin{equation}\label{eq:recurtime}
    \fpt\dg = \Lb\D.
  \end{equation}
  %
\end{thm}
\begin{proof}
  Pre-multiply \eqref{eq:fptrecdg} by $\eq\Lb\inv$:
  %
  \begin{equation*}
    \eq\Lb\inv\fpt\dg = \onev\trans,
  \end{equation*}
  %
  or in component form
  %
  \begin{equation*}
    \eq_i \Lb_{ii}\inv \fpt\dg_{ii} = 1.
  \end{equation*}
  %
  This is the same as \eqref{eq:recurtime} in component form: $\fpt\dg_{ii} = \Lb_{ii}/\eq_i$.
\end{proof}
I think the extra factor of $\Lb_{ii}$, compared to the discrete case \cite[Th.4.4.5]{kemeny1960finite}, occurs because in this case we are demanding that the process leaves the initial state once before returning, whereas in the discrete case we only measure the time it takes to go to the initial state after the first time-step.

We can substitute this into \eqref{eq:fptrecdg} to get a recursion relation for the off diagonal part:
%
\begin{equation}\label{eq:fptbrec}
  \MM\fptb = \D-\onem.
\end{equation}
%
\begin{thm}
  The off-diagonal solution of \eqref{eq:fptbrec} is unique.
\end{thm}
\begin{proof}
  Suppose we have two solutions for the off-diagonal part, $\fptb$ and $\fptb'$.
  Subtracting \eqref{eq:fptbrec} for these two solutions gives
  %
  \begin{equation*}
    \MM(\fptb-\fptb')=0.
  \end{equation*}
  %
  As we are assuming an ergodic process, the zero eigenvalue is non-degenerate.
  Therefore
  %
  \begin{equation*}
    \fptb-\fptb' = \onev\mathbf{y},
  \end{equation*}
  %
  for some row-vector $\mathbf{y}$. Looking at the diagonal components
  %
  \begin{equation*}
    \mathbf{y}_{i} = \fptb_{ii}-\fptb'_{ii} = 0,
  \end{equation*}
  %
  and therefore $\fptb-\fptb'=0$.
\end{proof}

\begin{thm}
  The off-diagonal mean first passage times are given by
  %
  \begin{equation}\label{eq:fptfund}
    \fptb = (\onem\fund\dg - \fund)\D.
  \end{equation}
  %
\end{thm}
\begin{proof}
  We just need to show that this satisfies \eqref{eq:fptbrec} and that the diagonal elements vanish.
  Premultiply \eqref{eq:fptfund} by $\MM$:
  %
  \begin{equation*}
  \begin{aligned}
    \MM\fptb &= -\MM\fund\D\\
     &= (\I-\onev\eq)\D \\
     &= \D-\onem,
  \end{aligned}
  \end{equation*}
  %
  where we used \eqref{eq:fundQZ} and \eqref{eq:pdotD}.
  The vanishing of the diagonal elements follows trivially from the component form of \eqref{eq:fptfund}:
%
\begin{equation}\label{eq:fptfundcmpt}
  \fptb_{ij} = \frac{\fund_{jj}-\fund_{ij}}{\eq_j}.
\end{equation}
%

\end{proof}

\subsection{Mixing time (Kemeny's constant)}\label{sec:mixtime}

\begin{thm}\label{th:kemenyconst}
  The quantity
  %
  \begin{equation}\label{eq:mixdef}
    \eta \equiv \sum_j \fptb_{ij}\eq_j
  \end{equation}
  %
  is independent of $i$.
\end{thm}
\begin{proof}
  We use \eqref{eq:fptfund}, \eqref{eq:fundrowsum} and the transpose of \eqref{eq:pdotD}:
  %
  \begin{equation*}
    \begin{aligned}
      \fptb (\eq)\trans &= (\onem\fund\dg - \fund) \D (\eq)\trans \\
        &= (\onev\onev\trans\fund\dg - \fund) \onev \\
        &= (\onev\trans\fund\dg\onev)\onev - \fund\onev \\
        &= (\tr\fund - \tau)\onev.
    \end{aligned}
  \end{equation*}
  %
  which proves \eqref{eq:mixdef} with $\eta = \tr\fund - \tau$.
\end{proof}

Note that it is essential that we use $\fptb$ and not $\fpt$ here, as that would lead to $\eta_i=\eta+\Lb_{ii}$, unlike the discrete time version \cite{hunter2006mixing} where this would only shift $\eta$ by 1.

\subsection{Sensitivity of equilibrium distribution}\label{sec:sensitivity}

Suppose that the Markov process, defined by $\MM$, depends on some parameter $\alpha$.
Differentiating \eqref{eq:funddef} gives
%
\begin{equation}\label{eq:diffZ}
  \diff{\fund}{\alpha} = \fund \diff{\MM}{\alpha} \fund.
\end{equation}
%
We can substitute this into the derivative of \eqref{eq:fundprob}:
%
\begin{equation}\label{eq:diffp}
  \diff{\eq}{\alpha} = \pib \fund \diff{\MM}{\alpha} \fund = \eq \diff{\MM}{\alpha} \fund.
\end{equation}
%
We can rewrite this in component form and use the fact that $\MM_{ii} = - \sum_{i\neq j} \MM_{ij}$:
%
\begin{equation}\label{eq:diffpT}
\begin{aligned}
  \diff{\eq_k}{\alpha} &= \sum_{i,j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} \\
    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} \fund_{jk} + \sum_i \eq_i \diff{\MM_{ii}}{\alpha} \fund_{ik} \\
    &= \sum_{i\neq j} \eq_i \diff{\MM_{ij}}{\alpha} (\fund_{jk} - \fund_{ik}) \\
    &= \sum_{i\neq j} \diff{\MM_{ij}}{\alpha} \eq_i \eq_k (\fptb_{ik} - \fptb_{jk}),
\end{aligned}
\end{equation}
%
which is the result of \cite{cho2000markov} that we need.
Note that the summand vanishes for $i=j$, so we can drop the restriction $i\neq j$ from the range of the sum.

\subsection{Subsets and flux}\label{sec:subsets}

Let us denote the set of states by $\CS$.
Consider a subset $\CA\subset\CS$.
We can define a projection operator onto this subset:
%
\begin{equation}\label{eq:proj}
  \prn{\I^\CA}_{ij} =
    \begin{cases}
      1 &\text{if $i=j\in\CA$,}\\
      0 &\text{otherwise.}
    \end{cases}
\end{equation}
%
We will use superscripts/subscripts to denote projection onto/summation over a subset:
%
\begin{equation}\label{eq:projsum}
  \begin{aligned}
    \pib^\CA &=\pib\I^\CA, &
    \M^{\cdot\CA}&=\M\I^\CA, &
    \M^{\CA\cdot}&=\I^\CA\M, &
    \mathbf{x}^\CA &= \I^\CA\mathbf{x},
    \\
    \pib_\CA &=\pib\onev^\CA, &
    \M_{\cdot\CA}&=\M\onev^\CA, &
    \M_{\CA\cdot}&=\prn{\onev^\CA}\trans\M, &
    \mathbf{x}_\CA &= \prn{\onev^\CA}\trans\mathbf{x},
  \end{aligned}
\end{equation}
%
where $\pib$ is a row vector, $\CM$ is a matrix and $\mathbf{x}$ is a column vector.

We can define a flux matrix, a.k.a.\ ergodic flow:
%
\begin{equation}\label{eq:flux}
  \F = \D\inv\MM,
  \qquad
  \F_{ij} = \eq_i \MM_{ij}.
\end{equation}
%
This measures the flow of probability between states in the equilibrium distribution.
Detailed balance, a.k.a.\ reversibility, is equivalent to $\F=\F\trans$.

The flux between two subsets is a particularly useful quantity:
%
\begin{equation}\label{eq:subflux}
  \F_{\CA\CB} = {\eq}^\CA\MM\onev^\CB.
\end{equation}
%
One can show that
%
\begin{equation}\label{eq:compflux}
  \F_{\CA\CA\comp} = \F_{\CA\comp\CA} = -\F_{\CA\CA} = -\F_{\CA\comp\CA\comp}
\end{equation}
%
using $\prn{{\eq}^\CA + {\eq}^{\CA\comp}}\MM=0$ and $\MM\prn{{\onev}^\CA + {\onev}^{\CA\comp}}=0$.

\subsection{Lumpability}\label{sec:lump}

Suppose we have partitioned the states into disjoint subsets, $\brc{\CA_\alpha}$:
%
\begin{equation}\label{eq:partition}
  \bigcup_\alpha \CA_\alpha = \CS,
  \qquad
  \CA_\alpha \cap \CA_\beta = \delta_{\alpha\beta}\CA_\alpha.
\end{equation}
%
We will use $\alpha$ instead of $\CA_\alpha$ in superscripts and subscripts in the following.
The fact that these subsets are disjoint and exhaustive allows us to define the function
%
\begin{equation}\label{eq:whichsey}
  \sigma(i)=\alpha
  \qquad\means\qquad
  i\in\CA_\alpha.
\end{equation}
%

We can use this partition to define a new stochastic process associated with the original Markov chain.
At time $t$, if the state of the original process is $i$, the state of the new process is $\sigma(i)$.

One may ask if this new process is a Markov chain.
The answer is yes, if the original Markov chain has a property called lumpability \wrt the partition
(see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time):
%
\begin{equation}\label{eq:lump}
  \sigma(i)=\sigma(j)
  \quad\implies \quad
  \MM_{i\alpha}=\MM_{j\alpha} \equiv \sum_{k\in\CA_\alpha} \MM_{jk},
\end{equation}
%
\ie the total transition rate from some state to some subset is the same for all starting states within the same subset.
This common value is the transition rate for the new lumped Markov chain.

This can be rewritten with the aid of two matrices
%
\begin{equation}\label{eq:lumpmats}
  U_{\alpha i} = \frac{\delta_{\alpha\sigma(i)}}{\abs{\CA_\alpha}},
  \qquad
  V_{i\alpha} = \delta_{\sigma(i)\alpha}.
\end{equation}
%
Left multiplication by $U$ averages over subsets, right multiplication by $V$ sums over subsets.
For $U$, we chose the uniform measure in each subset. Any measure would work equally well, \eg one proportional to the equilibrium distribution:
%
\begin{equation}\label{eq:altlumpmats}
  U_{\alpha i} = \frac{{\eq_i}^\alpha}{\eq_\alpha}.
\end{equation}
%

One can show that $(UV)_{\alpha\beta}=\delta_{\alpha\beta}$.
The matrix $VU$ is also interesting.
It has a block diagonal structure, with each block corresponding to a subset.
Each block is a discrete-time ergodic Markov matrix (it is an independent trials process with probabilities given by the measure chosen for $U$).
This means that the right eigenvectors with eigenvalue 1 will be those that are constant in each subset:
%
\begin{equation}\label{eq:setconst}
  VU\mathbf{x}=\mathbf{x}
  \qquad\means\qquad
  \mathbf{x} = \sum_\alpha x_\alpha \onev^\alpha.
\end{equation}
%

This allows us to write the lumpability condition \eqref{eq:lump}, and the transition matrix for the lumped process compactly:
%
\begin{equation}\label{eq:lumpcompact}
  VU\MM V = \MM V,
  \qquad
  \widehat{\MM} = U\MM V.
\end{equation}
%
By induction, one can show that similar relations hold for all powers:
%
\begin{equation}\label{eq:lumppower}
  VU\MM^nV = \MM^nV,
  \qquad
  \widehat{\MM}^n = U\MM^n V,
\end{equation}
%
and, via the Taylor series, for the exponential as well:
%
\begin{equation}\label{eq:lumpexp}
  VU \e^{t\MM} V = \e^{t\MM}V,
  \qquad
  \e^{t\widehat{\MM}} = U\e^{t\MM} V.
\end{equation}
%
The equilibrium distribution of the lumped process is given by
%
\begin{equation}\label{eq:lumpeq}
  \widehat{\mathbf{p}}^\infty = \eq V.
\end{equation}
%

\section{Signal-to-Noise ratio (SNR)}\label{sec:SNR}

In this section we will look at the signal-to-noise curve, and put an upper bound on its initial value.
We will only consider ergodic Markov chains.
Transient states would be unoccupied in equilibrium and would not be accessed by the signal creation process, therefore they could be removed from the analysis.
Absorbing chains are degenerate cases: they have zero initial signal but infinite decay times, so they can only be approached as the limit of a sequence of ergodic chains.

\subsection{Framework}\label{sec:framework}

The individual potentiation/depression events will be described by \emph{discrete}-time Markov chains:
%
\begin{equation}\label{eq:MWdef}
  \M^\pm \equiv \I + \W^\pm,
  \qquad
  \M^\pm\onev = \onev,
  \qquad
  \M^\pm_{ij} \in [0,1].
\end{equation}
%
The initial signal creation event occurs at time $t=0$, but all subsequent potentiation/depression events occur at random times according to Poisson processes with rates $rf^\pm$, where $f^++f^-=1$ are the fraction of plasticity events that are potentiating/depressing respectively.
This means that the ``forgetting'' process will be described by the \emph{continuous}-time Markov chain:
%
\begin{equation}\label{eq:forgetting}
  \MM = r \frg \equiv r\prn{f^+\W^+ + f^-\W^-}.
\end{equation}
%
We only require that this Markov chain is ergodic. The Markov chains described by $\M^\pm$ need not be.

We assume that the probability distribution starts in the equilibrium distribution \eqref{eq:equilibrium}.
During the initial signal creation, a fraction $f^+$ will change to $\eq\M^+$ and a fraction $f^-$ will change to $\eq\M^-$.
After this, probabilities will evolve according to \eqref{eq:rowsum}.

\subsection{SNR curve}\label{sec:SNRcurve}

As we ignore correlations between different synapses, the only quantity of interest will be the number of synapses with each strength.
The signal can be defined as the increase in synaptic strength due to potentiation plus the decrease in synaptic strength due to depression.
If there are $N$ synapses and the strength of the synapse in each state is given by the column vector, $\w$, this is given by
%
\begin{equation}\label{eq:signal}
  S(t) = N f^+ [\mathbf{p}(t|+)-\eq]\w - N f^+ [\mathbf{p}(t|-)-\eq]\w,
\end{equation}
%
where $\mathbf{p}(t|\pm)$ is the probability of being in each state conditional on potentiation/depression at $t=0$:
%
\begin{equation}\label{eq:initprob}
  \mathbf{p}(t|\pm) = \eq\M^\pm \e^{rt\frg}.
\end{equation}
%
We can rewrite the signal as
%
\begin{equation}\label{eq:signalr}
\begin{aligned}
  S(t) &= N \eq (f^+\W^+ - f^-\W^-)\e^{rt\frg}\w\\
    & N (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\frg}\w.
\end{aligned}
\end{equation}
%
For the noise, we can use the standard deviation of the synaptic strength:
%
\begin{equation}\label{eq:noise}
  \text{Noise} = \sqrt{N \prn{\sum_i \eq_i\w_i^2 - (\eq\w)^2}}.
\end{equation}
%
From now on we shall assume that there are only two possible synaptic strengths, strong and weak, which we will describe by $\w_i=\pm1$. This splits the set of states into two subsets:
%
\begin{equation}\label{eq:plusminusstates}
  \boldsymbol{+} = \set{k}{\w_k=+1},
  \qquad
  \boldsymbol{-} = \set{k}{\w_k=-1}.
\end{equation}
%
This simplifies the noise:
%
\begin{equation}\label{eq:noisepm}
  \text{Noise} = \sqrt{N \prn{4\eq_+\eq_-}}.
\end{equation}
%
The factors of $\eq_+\eq_-$ will have little effect on anything that follows, so we will drop them.\footnote{They will not affect the optimal transition rates for most quantities, and the optima will correspond to the symmetric case when $\eq_+=\eq_-=\half$.} This leaves the signal-to-noise ratio as
%
\begin{equation}\label{eq:SNRcurve}
  \SNR(t) = \sqrt{N} (2f^+f^-) \eq (\W^+ - \W^-)\e^{rt\frg}\w.
\end{equation}
%

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{?.tex} with its contents
%       or include ?.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%equilibrium distribution
\newcommand{\eq}{\mathbf{p}^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\uv}{\mathbf{u}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\enc}{\mathbf{q}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\syn}{\vec{w}}
\newcommand{\synid}{\syn_\text{ideal}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\DeclareMathOperator{\env}{Env}
\newcommand{\rh}{\hat{r}}
\newcommand{\CI}{\mathcal{I}}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
% ----------------------------------------------------------------
\input{sl_theorems_preamble.tex}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Laplacian envelope}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

@@

\begin{abstract}
  We try to find the continuous time Markov process that has the maximal Laplace tranformed signal-to-noise curve.
\end{abstract}

@@

\section{Recognition memory}\label{sec:recog}

We will be trying to store patterns in a set od $N$ synaptic weights, $\syn$.
Every time we try to store a pattern, these synapses are subjected to a plasticity event where each synapse is either potentiated or depressed, depending on the pattern.
we will assume that these patterns are spatially and temporally independent.

At some time, suppose we wish to determine if a given pattern is one of those that we previously attempted to store.
We wish to answer this question by looking at the synaptic weights directly (ideal observer).
For that given pattern there will be an ideal set of synaptic weights, $\synid$, where those synapses that were supposed to be potentiated are maximised and those that were supposed to be depressed are minimised.
Suppose that the given pattern was actually seen at time 0 and we are observing the synapses at time $t$.
The actual set of synaptic weights we see, $\syn(t)$, will be a vector of random variables that differs from $\synid$ due to the stochasticity of the pattern encoding and all of the other (uncorrelated) pattern that are stored after it.
As $t\to\infty$, the synaptic weights will become independent of the patter stored at $t=0$.
Thus, the vector of random variables $\syn(\infty)$ also describes the synaptic weights under the null hypothesis -- if that given pattern had never been stored.

We can test if the pattern had been previously stored by computing $\synid\cdt\syn$ and comparing it to some threshold.
For large $N$, this quantity will have a Gausssian distribution.
There will be a ROC curve as a function of this threshold:
%
\begin{equation}\label{eq:ROC}
  \begin{aligned}
  \operatorname{TPR} &= \Phi_c \prn{ \frac{ \Phi_c\inv(\operatorname{FPR}) - \snr }{ \operatorname{NNR} } },
  \quad\text{where } &
    \Phi_c(x) &= \int_x^\infty \frac{ \e^{-\frac{z^2}{2}} }{ \sqrt{2\pi} }\, \dz, \\&&
    \snr &= \frac{ \av{\synid\cdt\syn(t)} - \av{\synid\cdt\syn(\infty)} }{ \sqrt{\var(\synid\cdt\syn(\infty))}}, \\&&
    \operatorname{NNR} &= \sqrt{\frac{ \var(\synid\cdt\syn(t)) }{ \var(\synid\cdt\syn(\infty)) }},
  \end{aligned}
\end{equation}
%
and TPR/FPR are the true/false positive rates.
When the signal-to-noise ratio, $\snr$ is larger than $\Phi_c\inv(\operatorname{FPR})$, it is beneficial to decrease the noise-to-noise ratio, NNR.
In the other case, it is beneficial to increase it.
The expectations and variances are over the probability distribution of the synaptic states given the sequence of plasticity events and the probability distribution of the sequence of plasticity events.

The formulae above assumed that we know the time between storage and recognition.
We should also compute the expectations and variances over probability distribution of the recall time, $t$, as well.
If we only know average time, $\tau$, a natural choice for this distribution is
%
\begin{equation}\label{eq:recogtime}
  P(t\vert\tau) = \frac{\e^{-t/\tau}}{\tau}.
\end{equation}
%
Different parts of the brain, that store memories for different timescales, could be characterised by different values of $\tau$.
